Report on the Fourth Workshop on Sustainable Software for Science: Practice and Experiences (WSSSPE4)

### 2017-03-29 WSSSPE4_report/WSSSPE4_report.tex
Adding my affiliation details
Added final sentence and completed hanging sentence.

%
\thanks{{}$^{(1)}$ \hspace{-1ex}National Center for Supercomputing Applications (NCSA) \&
Department of Computer Science  \&
Department of Electrical and Computer Engineering  \&
School of Information Sciences (iSchool),
University of Illinois, Urbana--Champaign, IL, USA; d.katz@ieee.org; ORCID: 0000-0001-5934-7525}
%
\thanks{{}$^{(2)}$ School of Mechanical, Industrial, and Manufacturing Engineering,
Oregon State University, Corvallis, OR, USA; kyle.niemeyer@oregonstate.edu; ORCID: 0000-0003-4425-7097}
%
\thanks{{}$^{(3)}$ Center for Research Computing \& Department of Computer Science and Engineering,
University of Notre Dame, IN, USA; sandra.gesing@nd.edu; ORCID: 0000-0002-6051-0673}
%
\thanks{{}$^{(4)}$ University of California, Davis, CA, USA; ljhwang@ucdavis.edu; ORCID: ?}
%
\thanks{{}$^{(5)}$ \note{info needed}}
%
\thanks{{}$^{(6)}$ Astrophysics Source Code Library, Astronomy Department, University of Maryland, College Park, MD, USA; aallen@ascl.net; ORCID: 0000-0003-3477-2845}
%
\thanks{{}$^{(7)}$ \note{info needed}}
%
\thanks{{}$^{(8)}$ \note{info needed}}
%
\thanks{{}$^{(9)}$ \note{info needed}}
%
\thanks{{}$^{(10)}$ \note{info needed}}
%
\thanks{{}$^{(11)}$ University of California, Berkeley, Berkeley Institute for Data Science (BIDS), Berkeley, CA, USA; stuart@stuartgeiger.com; ORCID: 0000-0001-7215-0532}
%
\thanks{{}$^{(12)}$ \note{info needed}}
%
\thanks{{}$^{(13)}$ \note{info needed}}
%
\thanks{{}$^{(14)}$ Department of Mathematics, Colorado State
  University, Fort Collins, CO 80523, USA; bangerth@colostate.edu; ORCID: 0000-0003-2311-9402}

@@

To investigate best practices for scientific software we put together a panel of five people with different perspectives on ``Best Practices for Scientific Software''.
Alice Allen, Editor, Astrophysics Source Code Library, brought an understanding of the difficulties of organizing a community and curating their software.
Mike Croucher from the University of Sheffield and Rob Haines from the University of Manchester are both Research Software Engineers with decades of experience of writing code for researchers.
Patricia Lago from Vrije Universiteit Amsterdam presented a keynote at the workshop (\S\ref{sec:keynote}) and brought a fresh perspective on software sustainability from the point of view of its impact on society and business.
Finally, Karthik Ram from the Berkeley Institute for Data Science brought his perspective on the practicalities of using scientific software to conduct his research as a data scientist.

The panel opened with a question to the floor ``Who in the audience still codes in anger?'' which caused some confusion as, it turns out, that this is a turn of phrase that does not span the Atlantic.
After translating this into the more open ``Who in the audience still writes code as part of their day-to-day work?'' we were able to confirm the suspicion that the WSSSPE audience is heavily involved in, not just using, but also writing code.
Hence, the following discussions would \todo{what goes here?}

The panel opened by noting that the world of scientific computing is changing rapidly, and the panel asked which practices have emerged recently that, if not handled correctly, might be the next generation's problem?
This quickly led to a discussion about the dangers of containerization.
These technologies are pushed by some---a minority it is hoped---as a panacea for sustainability.
The panel and attendees generally agreed that containerization is useful, but should not be seen as the solution to sustainability.
After all, these technologies are just as liable as any to decay.

A later question looked at the common practices from software engineering that the scientific community should be using.
The discussion that followed picked up on the key elements of sustainability, version control, documentation and ensuring that communities are active and engaged.
When asked whether there was resistance from the research community to adopting these practices, few negative experiences were reported.
It would appear that the problem lies in getting the message out about these technologies rather than trying sell researchers on the benefit of them.

The panel rounded up with a final question about "If you wanted to have the most disastrous effect on scientific software, what one practice would you advise people to keep doing?

\todo{} <WILL ADD HERE ONCE I'VE TALKED TO PEOPLE>

>>>>

%
\thanks{{}$^{(1)}$ \hspace{-1ex}National Center for Supercomputing Applications (NCSA) \&
Department of Computer Science  \&
Department of Electrical and Computer Engineering  \&
School of Information Sciences (iSchool),
University of Illinois, Urbana--Champaign, IL, USA; d.katz@ieee.org; ORCID: 0000-0001-5934-7525}
%
\thanks{{}$^{(2)}$ School of Mechanical, Industrial, and Manufacturing Engineering,
Oregon State University, Corvallis, OR, USA; kyle.niemeyer@oregonstate.edu; ORCID: 0000-0003-4425-7097}
%
\thanks{{}$^{(3)}$ Center for Research Computing \& Department of Computer Science and Engineering,
University of Notre Dame, IN, USA; sandra.gesing@nd.edu; ORCID: 0000-0002-6051-0673}
%
\thanks{{}$^{(4)}$ University of California, Davis, CA, USA; ljhwang@ucdavis.edu; ORCID: ?}
%
\thanks{{}$^{(5)}$ The Software Sustainability Institute, Southampton, UK; s.hettrick@software.ac.uk; ORCID: 0000-0002-6809-5195}
%
\thanks{{}$^{(6)}$ Astrophysics Source Code Library, Astronomy Department, University of Maryland, College Park, MD, USA; aallen@ascl.net; ORCID: 0000-0003-3477-2845}
%
\thanks{{}$^{(7)}$ \note{info needed}}
%
\thanks{{}$^{(8)}$ \note{info needed}}
%
\thanks{{}$^{(9)}$ \note{info needed}}
%
\thanks{{}$^{(10)}$ \note{info needed}}
%
\thanks{{}$^{(11)}$ University of California, Berkeley, Berkeley Institute for Data Science (BIDS), Berkeley, CA, USA; stuart@stuartgeiger.com; ORCID: 0000-0001-7215-0532}
%
\thanks{{}$^{(12)}$ \note{info needed}}
%
\thanks{{}$^{(13)}$ \note{info needed}}
%
\thanks{{}$^{(14)}$ Department of Mathematics, Colorado State
  University, Fort Collins, CO 80523, USA; bangerth@colostate.edu; ORCID: 0000-0003-2311-9402}

@@

To investigate best practices for scientific software we brought together a panel of five experts with different perspectives on ``Best Practices for Scientific Software''. These were: Alice Allen, Editor, Astrophysics Source Code Library, brought an understanding of the difficulties of organizing a community and curating their software. Mike Croucher from the University of Sheffield and Rob Haines from the University of Manchester who are both Research Software Engineers with decades of experience of writing code for researchers. Patricia Lago from Vrije Universiteit Amsterdam presented a keynote at the workshop (\S\ref{sec:keynote}) and brought a fresh perspective on software sustainability from the point of view of its impact on society and business. Finally, Karthik Ram from the Berkeley Institute for Data Science brought his perspective on the practicalities of using scientific software to conduct his research as a data scientist.

The panel opened with a question to the floor ``Who in the audience still codes in anger?'' which caused some confusion as, it turns out, that this is a turn of phrase that has not crossed the Atlantic. After translating this question into the more open ``Who in the audience still writes code as part of their day-to-day work?'' we were able to confirm the suspicion that the WSSSPE audience is heavily involved writing, not just using, code. Hence, the panel were going to have to deal with questions from a highly experienced audience.

The panel opened by noting that the world of scientific computing is changing rapidly. Each panelist was asked for their views on practices which had recently emerged and, if not handled correctly, would become the basis for the next generation's software problems?
This quickly led to a discussion about the dangers of containerization. These technologies are pushed by some---a minority it is hoped---as a panacea for software sustainability. The panel and attendees generally agreed that containerization is useful, but should not be seen as the solution to sustainability. After all, these technologies are just as liable as any to decay.

A later question looked at the common practices from software engineering that the scientific community should be using. The discussion that followed picked up on the key elements of basic sustainable practice: version control, documentation and ensuring that communities are active and engaged. When asked whether there was resistance from the research community to adopting these practices, few negative experiences were reported. From the panel's point of view, it would appear that the problem in getting researchers to employ sustainable practicies lies more in getting the message out to researchers, rather than selling them on the benefits.

The panel rounded up with a final question about "If you wanted to have the most disastrous effect on scientific software, what one practice would you advise people to keep doing?" A few good responses were discussed, but the biggest laugh went to Mike Croucher who said "Containerize everything!".
